Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-15 methods.tex
minor edits to methods
adding code and packages to compile code
edited the code listing

\col is a simple C program that repeatedly initializes the upper
triangle of a 2048 $\times$ 2048 matrix in column-major order.

{\color{red}Joshua, can you please insert code here?}

\noindent As mentioned in Section~\ref{sec:intro}, this simple program
exhibits surprisingly complicated behavior~\cite{mytkowicz09}.  A time
series of the processor performance during the execution this program
is shown in Figure~\ref{fig:col-ts}. \alert{would it be useful to
  include said 3-line loop?}
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\columnwidth]{figs/colFullTS}
    \caption{The instructions executed per CPU clock cycle (IPC) as
      the \col program runs.  Each point is the average IPC in a
      100,000 instruction period. Figure~\ref{fig:col-ipc} is a
      closeup of the first part of this signal.  }
    \label{fig:col-ts}

>>>>

\col is a simple C program that repeatedly initializes the upper
triangle of a 2048 $\times$ 2048 matrix in column-major order by
looping over the following three lines of code:
\begin{lstlisting}
% #define SIZE (2048)
% 
% int main(int argc, char ** argv)
% {
%   int i, j, n;
%   static int data[SIZE][SIZE];
%   
%   for(n=0;n<numInits;n++){  
  for(i=0;i<SIZE;i++)
  	for(j=i;j<SIZE;j++)
  		data[j][i] = 0;
%   }
%   return 0;
}
\end{lstlisting}
\noindent As mentioned in Section~\ref{sec:intro}, this simple program
exhibits surprisingly complicated behavior~\cite{mytkowicz09}.  A time
series of the processor performance during the execution of this
program is shown in Figure~\ref{fig:col-ts}.
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\columnwidth]{figs/colFullTS}
    \caption{The instructions executed per CPU clock cycle (IPC) as
      the \col program runs.  Each point is the average IPC in a
      100,000 instruction period. Figure~\ref{fig:col-ipc} is a
      closeup of the first part of this signal.  }
    \label{fig:col-ts}
